---
title: "The Beta distribution Quiz Frederic Brieger"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    social: menu

---

```{r setup, include=FALSE}
library(flexdashboard)
```

### Question 1: Why should we use the beta distribution and to what distribution is it related?

```{r}
val <- seq(0, 1, 0.01) # Spacing
par(mfrow = c(1,1))
plot(val, dbeta(val, 3, 3), cex.lab = 1.5, type = "l", lwd = 2, main = "a, b between 1 and 3", ylab = "dbeta")
for (a in 1:3){
  for (b in 1:3){
  lines(val, dbeta(val, a, b), cex.lab = 1.5, type = "l", lwd = 2, col = rgb(runif(1),runif(1),runif(1),1))}}

```

***

The Beta distribution is a discrete distribution that looks similar to the binomial distribution. When applied to the interval [0, 1] it can be used to describe a prior or posterior probability distribution. For that, we can manipulate the values a and b or work just from a most probable value.

In this plot you can see how the pdf is affected by different a and b values (between 1 and 3).


### Question 2: What could we use the beta distribution for in the field of natural hazards?

```{r}

# Using the MODE of our prior  distribution
gesamt <- 36
w <- 0.7
a <- w*(gesamt - 2) + 1
b <- (1-w) * (gesamt - 2) + 1

plot(val, dbeta(val, a, b), cex.lab = 1.5, type = "l", lwd = 2, col = "firebrick1")
print(c(a,b))
```

*** 

The beta distribution is the perfect choice if we are looking for a prior distribution with a distinct maximum.
For example: An old paper suggests that most wildfires out of 36 events reduce the vegetation by 70 %, but the data this value is based on is missing.
Given this MODE we can plot a prior probability distribution using the beta distribution.

### Question 3: How can we use the mean or mode to calculate a beta distribution?


```{r, echo = FALSE}
val <- seq(0, 1, 0.01) # Spacing
par(mfrow=c(2,1)) # Plotlayout

# Using the MEAN of our prior  distribution
gesamt <- 10
m <- 0.5
a <- m * (gesamt)
b <-  (a/m)-a

plot(val, dbeta(val, a, b), cex.lab = 1.5, type = "l", lwd = 2, col = "gold1", ylab = "dbeta")
print(paste("a1 =", a, ", b1 =", b))

# Using the MODE of our prior  distribution
gesamt <- 100
w <- 0.8
a <- w*(gesamt - 2) + 1
b <- (1-w) * (gesamt - 2) + 1

plot(val, dbeta(val, a, b), cex.lab = 1.5, type = "l", lwd = 2, col = "dodgerblue2", ylab = "dbeta")
print(paste("a2 =", a, ", b2 =", b))


```
***
```{r, echo = TRUE, eval = FALSE}
val <- seq(0, 1, 0.01) # Spacing
par(mfrow=c(2,1)) # Plotlayout

# Using the MEAN of our prior  distribution
gesamt <- 10
m <- 0.5
a <- m * (gesamt)
b <-  (a/m)-a

plot(val, dbeta(val, a, b), cex.lab = 1.5, type = "l", lwd = 2)
print(c(a,b))

# Using the MODE of our prior  distribution
gesamt <- 100
w <- 0.8
a <- w*(gesamt - 2) + 1
b <- (1-w) * (gesamt - 2) + 1

plot(val, dbeta(val, a, b), cex.lab = 1.5, type = "l", lwd = 2)
print(c(a,b))

```
