---
title: "Gauss homework Frederic"
author: "Frederic Brieger"
date: "22 Januar 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = TRUE}

```
## Where does the Gaussian distribution come from?

The gaussian or normal distribution is a continuous probability distribution and has a symmetrical shape. Both sides of the symmetry axis resemble a simple exponential function.

```{r, echo = TRUE}
x <- seq(1, 10, 0.1)
y <- exp(-x)

plot(x,y, type = "l", lwd = 2, col = "firebrick1", main = "Basic exponential function", ylab = "exp(-x)")
grid()

```


## How does the normal distribution look like?

The distribution can be described with two parameters: the mean, mode and median of the distribution is mu and the standart deviation is sigma.
$f(x| mu, sigma^2) = 1/(sqrt(2*pi*sigma^2) * exp(-((x-mu)^2 / (2*sigma^2))$

```{r, echo = TRUE}
x2 <- seq(-5, 5, 0.1)
plot(x2, dnorm(x2, 0, 1), type = "l", lwd = 2, col = "firebrick1", main = "Mu = 0, sigma = 1", ylab = "p(x), cumm. p(x)", ylim = c(0, 1)) # Plotting the desnsity function
grid()
lines(x2, pnorm(x2, 0, 1), type = "l", lwd = 2, col = "deepskyblue1") # Adding the distribution function
legend(-4.5, 1, legend = c("Density function", "Distribution function"), col = c("firebrick1", "deepskyblue1"), lty = 1)
```

To see how sigma and mu change the shape of the density- and its distribution function I created a 3x3 plot with variing values.

```{r, echo = TRUE}
x3 <- seq(-10, 10, 0.1)
mus <- c(1,2,3) # Different mus
sigmas <- c(1,2,3) # Different sigmas

par(mfrow=c(3,3)) # 3x3 plot matrix
for (i in 1:3){
  for (t in 1:3){
    plot(x3, dnorm(x3, mean = mus[i], sd = sigmas[t]), type = "l", 
         lwd = 2, col = "turquoise3", ylab = "p(x)", xlim = c(-10, 10), 
         ylim = c(0, 0.4), main = paste("mu =", i, ", sd =", t))
    grid()
  }
}
```

We can see that mu does not change the form of the function but moves it along the x-axis. The standard deviation sigma changes the form of the bell.
As it is a probability distribution function, the area under the graph always remains 1. 

## Typical applications of the normal distribution

The importance of the Gaussian distribution comes from the fact that many natural processes follow this distribution such as errors from measurements, grades from an exam or molecule motion.

## Usage in the field of natural hazards or data analysis

As we learned, many natural processes follow the normal distribution and so can data which is derived from these processes. 
In this example a dataset containing five normally distributed values needs to be testet if it is originated from a datastation with a given standart deviation and long term mean value.

```{r, echo = TRUE}
sigma <- 5
data <- rnorm(5, runif(1, 50, 70), sigma) # Generates a normally distributed dataset containing 5 numbers
testmu <- runif(1, 50, 70) # A mu for wich I want to test the data
likelihood <- prod(dnorm(data, testmu, sigma)) # Likelihood p(testmu|data)

print(data)
print(paste("testmu=", testmu))
print(paste("Likelihood", likelihood))
```
