---
title: "Hausaufgabe 3"
author: "Kuras, Bonitz"
date: "7 Dezember 2017"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```





```{r tabsets, echo=FALSE}
library(shiny)

ui <- fluidPage(
  titlePanel("Bernoulli Application"),
  sidebarPanel(
      numericInput(inputId = "success","Success",
               value = 1,
               min = 1),
      numericInput(inputId = "failure","Failure",
               value = 1,
               min = 1),
      numericInput(inputId = "numprior", "Number of Priors",
                   value = 10,
                   min = 1)
  ),
 mainPanel(tabsetPanel(
  tabPanel(
      title = "Graphs",
      plotOutput("plots")
     
    )
  ))
)

server <- function(input, output) {
# Anzahl des Prior Inputs ausgeben
  theta <- reactive({
    numprior <- input$numprior
    theta <- seq(0,1,1/(numprior+1))
    theta <- theta[2:(length(theta)-1)]
    })
#Prior Verteilung
  prior <- reactive({
      prior <- pmin(theta(), 1-theta())
      prior <- prior/sum(prior)
    })
  
#berechnung der likelihood
  likelihood <- reactive(choose((input$success+input$failure),input$success)*(theta()^input$success*(1-theta())^input$failure))
 
#berechnung des posterior
  posterior <- reactive((likelihood()*prior())/((likelihood()*prior())+((1-likelihood())*(1-prior())))) 
 
  
  output$plots <- renderPlot({
    layout(matrix(c(1, 2, 3), nrow = 3, ncol = 1, byrow = FALSE))
    par(mai = c(0.5, 0.5, 0.5, 0.1))
    
    plot(theta(),prior(),
         type = "h",
         main = "Prior",
         xlab = "Theta",
         ylab = "Prior",
         xlim = c(0,1))
    plot(theta(),likelihood(),
         type = "h",
         main = "Likelihood",
         xlab = "Theta",
         ylab = "Likelihood",
         xlim = c(0,1))
    plot(theta(),posterior(),
         type = "h",
         main = "Posterior",
         xlab = "Theta",
         ylab = "Posterior",
         xlim = c(0,1))
    

})
}

shinyApp(ui, server)